MLISA: Machine Learning-based \\ Intelligent Service Agents

### 2014-07-12 papers/webservice-coalition-learning/webservice_coalition_learning.tex
changesb

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{TO CHANGE: Web Service Community \\ Learning}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

@@

\IEEEPARstart{O}{ver} the past years, online services have become important part of many scalable business applications. The increasing reliance on online service providers has significantly influenced the way web services are engineered. Therefore, given the dynamic and unpredictable nature of the Internet, delivering high quality services is still a critical and challenging issue. One approach to deliver high quality service is utilizing intelligent decision making agents that are aimed at maximizing gain by exploring best ways of providing services that satisfy end users. The recent developments attempted to shift web services from passive model in the form of individual components to autonomous and group-based components that follow common goals. In group-based models, interaction, composition, and cooperation are the key challenges \cite{ICWS2011-1, SCC2011-1, journals/mags/BaldoniBM10, journals/jcss/CasadoYT13} that have direct influence on group's overall performance in achieving common goals. All that being said, the idea of \emph{community} of web services has been emerged that gathers services having similar functionality but distinct non-functional properties \cite{Zeng:2003:QDW:775152.775211, Paik:2005:TSS:2229263.2230038, Medjahed05adynamic, 10.1109/ARES.2008.7}. A \emph{community} of web service runs continuous performance assessment functions that regulate web services' interactions and manage their composition and cooperation.

The advantages of community of Web service is that they facilitate the discovery of web services and as community, they provide better quality services. In other words, communities act as abstract web services, communicating with external entities via same standard protocols that a normal web services employs. The down side to communities is the management complexity that involves finding/inviting individual web services and managing the overall quality as combined work done by web services that although have similar functionality, but have different attitudes. Interacting with communities of web services, users send their requests to the coordinators of the community, which plays the role of the community representative or access point. The community coordinator in \emph{community centralized} architecture is responsible for task recieve/delivery issues. Moreover, as community representative, it verifies credentials of new web services before accepting them to join the community as well as firing web services, which do not add enough value to the community.

Recent work have emphasized on motivation and benefits of communities of web services [references]. This is due to the fact that communities facilitate discovery of web services, improve the quality of service, and help individual services find better market share and opportunities. The challenge is then community development, which is finding the right web service for the community, satisfying all web services and community coordinators involved within the community. This is also a challenge from individual web service side in the sense that a web service seeks for communities to interact with and obtain best outcome. Towards addressing these issues, there have been a number of proposals. Some of these works propose complex models on demand algorithms \cite{DBLP:conf/IEEEscc/LimTMB12, 10.1109/TSC.2012.12, 10.1109/TSC.2014.2312940} to effectively find target points. But they have to simplify the algorithm by limited parameters in order to make it feasible and algorithm executable. The simplifications problem some times negatively influences the outcome in the sense that simplifying the constraints may cause ignoring important aspects of the problem. For example \cite{10.1109/TSC.2012.12} proposes a simple egalitarian way of distributing gain, which completely ignores gain for collaborative work instead of complex shapely-value. This model considers gain for all subset of agents working together, which basically ignores the most important criteria of finding optimal group. Other category of related work exists which they do not focus on all parties involved within the community \cite{DBLP:conf/IEEEscc/KhosravifarABT11}. 
%And some other work which do not focus on rationality on web services or communities involved [refrences]. 
These solutions generally suffer from high complexity, which makes decision making impossible in an on-demand fashion, or they simplify important aspects to make it practical in real world hurting the decision making performance.

In this paper, we have implemented [OurMethodName] using which, we initially train web services in different possible communities and let them experiment and check the outcome of different scenarios with respect to  their internal functional and external cooperative characteristics. After being trained, web services get to compute expectations as utilities they would gain while cooperating with communities of different characteristics. In contrast, communities can consider the training into account and analyze the characteristics of different individual web services and make prudent decisions while inviting a web service to join or accepting a join inquiry initiated from a web service. In general, the trained results equip web services with efficient methods of foreseeing how their choices of actions would impact their long term and short term goals, therefore opting for best decision available.

In further details, the proposed learning method is applicable for any sort of intelligent agent-based model. To effectively train web services, we used real data set (information about the data set) to extract web services' individual characteristics and used them to measure outcomes when they cooperate with one another. Engineering these features, we gathered functional and cooperative features for both individual web services and communities in 60 different time slots. We were able to investigate the path a web service might take to achieve the best utility out of effective interaction with others. All paths and the outcomes are labeled to be utilized in training model. Using cross validation set, web services are able to compute optimal hypothesis function (using logistic regression) that can be used to predict outcome out of cooperative work with other individual web services or communities. Our findings show that web services equipped with this framework have by far better outcome in the form of cooperative service providers than the ones that either do not cooperate or randomly find communities to join.

@@

\section{The Model}\label{s:preliminaries}
In this section, we first present the architecture of the proposed model. We explore the characterestic of intelligent service agents and the way we train them. ...

@@

In our research work, we abstract web services as rational agents\footnote{The term rational is used here in the sense that web services are utility maximizers} providing services to end users. They aim to maximize their individual income  by receiving enough requests from end users. In order to increase their revenue, web services seek for more tasks if they have the capacity and throughput to do so. Web services can join communities to have better efficiency by collaborating with others, to have access to broad market share, and to have opportunity of receiving a bigger task pool from end users. Furthermore, the high reliance on web services has increased quality expectations from end users. Communities of web services can provide higher availability, performance, reliability, and recovery for end users.

@@

With a group of web services having identical or similar functionalists, quality of service metrics provide nonfunctional characteristics for the optimal candidate selection. Web service quality metrics have been presented in \cite{Ardagna:2007:ASC:1263152.1263531,Menasce:2002:QIW:613357.613758,10.1109/ISSRE.2011.17}, we have adopted the most representative QoS properties of web services which highly influence the utility of web services.

@@

  \item \emph{Throughput} is the rate at which a service can process requests. QoS measures can include the maximum throughput or a function that describes how throughput varies with load intensity.

	\item \emph{Availability} is the percentage of time that a service is operating.
	The probability that the service operation is accessible. Availability is a number in the range $[0, 1]$.

	\item \emph{Execution Time} is the time a service takes to respond to various types of requests. 
	%is the expected delay between the time instant when a request is sent and the time when the result is obtained. 
	Execution time is measured in seconds. Execution time can be affected by load intensity, which can be measured in terms of arrival rates (such as requests per second) or number of concurrent requests.

@@

\begin{itemize}
  \item \emph{External Parameter 1} $(exp_1)$. is an estimate of how close the web service's \emph{execution time} to other web services is. It is the difference between web service's \emph{response time} metric and the maximum value of response time of web services, that the web service has recently gathered information about.
	\item \emph{External Parameter 2} $(exp_2)$ is a comparison of the web service's rate of performing tasks to other competing web services. It is the difference between web service's \emph{throughput} metric and the maximum value of throughput of web services, that the web service has recently gathered information about.
\end{itemize}

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{MLISA: Machine Learning-based \\ Intelligent Service Agents}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

@@

\IEEEPARstart{O}{ver} the past years, online services have become important part of many scalable business applications. The increasing reliance on online service providers has significantly influenced the way web services are engineered. Therefore, given the dynamic and unpredictable nature of the Internet, delivering high quality services is still a critical and challenging issue. One practical solution that delivers high quality service is utilizing intelligent decision making agents. These agents are aimed at maximizing gain by exploring best ways of providing services that satisfy end users. However, agent-based web services are functionally limitted in the sense that they cannot handle large number of requests at the same time, otherwise their quality of provided service would diminish. The recent developments attempted to shift web services from passive model in the form of individual components to autonomous and group-based components that follow common goals. In group-based models, interaction, composition, and cooperation are the key challenges \cite{ICWS2011-1, SCC2011-1, journals/mags/BaldoniBM10, journals/jcss/CasadoYT13} that have direct impact on group's overall performance in achieving common goals. All that being said, the concept of \emph{community} of web services has been emerged that gathers services having similar functionality but distinct non-functional properties \cite{Zeng:2003:QDW:775152.775211, Paik:2005:TSS:2229263.2230038, Medjahed05adynamic, 10.1109/ARES.2008.7}. A \emph{community} of web service runs continuous performance assessment functions that regulate web services' interactions and manage their composition and cooperation.

The advantages of community of Web services is that the community facilitates the discovery of web services and as community provides better quality services compared to passive models with individual components. In other words, communities act as abstract web services, communicating with external entities via same standard protocols that a normal web services employs. But they requlate the service process via sophisticated internal communication protocols that result in providing a service out of combined work of a number of web services. The down side to communities is the management complexity that involves finding/inviting individual web services and managing the overall quality as combined work done by web services that although have similar functionality, but have different attitudes. Interacting with communities of web services, users send their requests to the coordinators of the community, which plays the role of the community representative or access point. The community coordinator in \emph{community centralized} architecture is responsible for task recieve/delivery issues. Moreover, as community representative, it verifies credentials of new web services before accepting them to join the community as well as firing web services, which do not add enough value to the community.

Recent work have emphasized on motivation and benefits of communities of web services [references]. This is due to the fact that communities facilitate discovery of web services, improve the Quality of Service (QoS), and help individual services find better market share and opportunities. Now there is a twofold challenge that is still not addressed: (1) the first challenge is community development, which is finding the right web service for the community, satisfying all web services and community coordinators involved within the community; and (2) the second challenge is from individual web service side in the sense that a web service seeks for communities to interact with and obtain best outcome. Towards addressing these research problems, there have been a number of proposals. Some of these works propose complex models on demand algorithms \cite{DBLP:conf/IEEEscc/LimTMB12, 10.1109/TSC.2012.12, 10.1109/TSC.2014.2312940} to effectively find target points. But they have to simplify the algorithm by limited parameters in order to make it feasible and algorithm executable. Simplified problems some times negatively influence the outcome in the sense that simplifying the constraints may cause ignoring important aspects of the problem. For example \cite{10.1109/TSC.2012.12} proposes a simple egalitarian way of distributing gain, which completely ignores gain for collaborative work instead of complex shapely-value. This model considers gain for all subset of agents working together, which basically ignores the most important criteria of finding optimal group. Other category of related work exists which they do not focus on all parties involved within the community \cite{DBLP:conf/IEEEscc/KhosravifarABT11}. 
%And some other work which do not focus on rationality on web services or communities involved [refrences]. 
These solutions generally suffer from high complexity, which makes decision making impossible in an on-demand fashion, or they simplify important aspects to make it practical in real world hurting the decision making performance.

In this paper, we introduce MLISA (Machine Learning-based Intelligent Service Agents) community framework that operates based on a trained model that regulates web service agents' decision making process in terms of cooperating with one another. This model uses the trained data in the form of information obtained from large number of web services regarding their single and cooperative utilities as well as environmental parameters such as demand, service quality, etc. The training model deploys logistic regression algorithm to build hypothesis function that can thoroughly address the aforementioned research problems. Our findings show that communities of web services can efficiently find the appropriate web service to invite cooperating with them as well as a single web service can find best communities to target joining them. In further details, after being trained, web services get to compute expectations as utilities they would gain while cooperating with communities of different characteristics. In contrast, communities can consider the training into account and analyze the characteristics of different individual web services and make prudent decisions while inviting a web service to join or accepting a join inquiry initiated from a web service. In general, MLISA equips web services with efficient methods of foreseeing how their choices of actions would impact their long-term and short-term goals, therefore opting for best decision available.

In further details, MLISA is applicable for any sort of intelligent agent-based model. To effectively train web services, we used real data set (information about the data set) to extract web services' individual characteristics and used them to measure outcomes when they cooperate with one another. Engineering and extracting these features, we gathered functional and cooperative features for both individual web services and communities in 60 different time slots. We were able to investigate the path a web service might take to achieve the best utility out of effective interaction with others. All paths and the outcomes are labeled to be utilized in training model. Using cross validation set, web services are able to compute optimal hypothesis function (using logistic regression) that can be used to predict outcome out of cooperative work with other individual web services or communities. Our findings show that web services equipped with MLISA have by far better outcome in the form of cooperative service providers than the ones that either do not cooperate or randomly find communities to join.

@@

\section{The Model}\label{s:preliminaries}
In this section, we first present the architecture of MLISA. We explore the characterestics of intelligent service agents and the features we extract for training. To do this, in the following, we discuss some preliminaries.

@@

In our research work, we abstract web services as rational agents\footnote{The term rational is used here in the sense that web services are utility maximizers} providing services to end users. They aim to maximize their individual income  by receiving enough requests from end users. In order to increase their revenues, web services seek for more tasks if they have the capacity and throughput to do so. Web services can join communities to have better efficiency by collaborating with others, to have access to broad market share, and to have opportunity of receiving a bigger task pool from end users. Furthermore, the high reliance on web services has increased quality expectations from end users. Communities of web services can provide higher availability, performance, reliability, and recovery for end users.

@@

With a group of web services having identical or similar functionalists, QoS metrics provide nonfunctional characteristics for the optimal candidate selection. Web service quality metrics have been presented in \cite{Ardagna:2007:ASC:1263152.1263531,Menasce:2002:QIW:613357.613758,10.1109/ISSRE.2011.17}, we have adopted the most representative QoS properties of web services which highly influence the utility of web services.

@@

  \item \emph{Throughput} is the rate at which a service can process requests. QoS measures can include the maximum throughput or a function that describes how throughput varies with load intensity. Throughput is a real number in the range $[0, 1]$.

	\item \emph{Availability} is the percentage of time that a service is operating.
	The probability that the service operation is accessible. Availability is also a real number in the range $[0, 1]$.

	\item \emph{Execution Time} is the time a service takes to respond to various types of requests. 
	%is the expected delay between the time instant when a request is sent and the time when the result is obtained. 
	Execution time is measured in seconds. Execution time can be affected by load intensity, which can be measured in terms of arrival rates (such as requests per second) or number of concurrent requests. This internal feature is a positive integer.

@@

\begin{itemize}
  \item \emph{External Parameter 1} $(exp_1)$. is an estimate of how close the web service's \emph{execution time} is to other web services. It is the difference between web service's \emph{response time} metric and the maximum value of response time of web services, that the web service has recently gathered information about.
	\item \emph{External Parameter 2} $(exp_2)$ is a comparison of the web service's rate of performing tasks to other competing web services. It is the difference between web service's \emph{throughput} metric and the maximum value of throughput of web services, that the web service has recently gathered information about.
\end{itemize}
